 

### 2020-09-21 vertical-thermal-review.Rmd
updated

Globally, temperatures have risen # $^\circ$C since YEAR and are expected to rise an additional #-#$^\circ$C by YEAR (IPCC REF). 
There is also increasing severity and frequency of heat waves (Perkins et al. 2012).
This is expected to have profound effects on forest biodiversity and ecosystem function (REFS), and the resulting feedbacks to the climate system (e.g., decreased C storage, altered albedo and hydrology) will have a critical influence on the future of trajectory of climate change [@bonan_ecological_2016, Bonan 2008].
With increases in temperature, temperate and tropical forest canopies are already or will be soon functioning beyond their optimal photosynthetic threshold [@mau_temperate_2018, Huang et al. 2019]. 
This, alongside climate change-driven disturbances such as drought, may contribute to increasing tree mortality in forests around the world [@mcdowellMechanismsPlantSurvival2008, @allenUnderestimationGlobalVulnerability2015, Brienen et al. 2015].

@@

>>>>

Globally, temperatures have risen # $^\circ$C since YEAR and are expected to rise an additional #-#$^\circ$C by YEAR (IPCC REF). 
There is also increasing severity and frequency of heat waves (@perkinsIncreasingFrequencyIntensity2012).
This is expected to have profound effects on forest biodiversity and ecosystem function (REFS), and the resulting feedbacks to the climate system (e.g., decreased C storage, altered albedo and hydrology) will have a critical influence on the future of trajectory of climate change [@bonan_ecological_2016, Bonan 2008].
With increases in temperature, temperate and tropical forest canopies are already or will be soon functioning beyond their optimal photosynthetic threshold [@mau_temperate_2018, @huangAirTemperatureOptima2019]. 
This, alongside climate change-driven disturbances such as drought, may contribute to increasing tree mortality in forests around the world [@mcdowellMechanismsPlantSurvival2008, @allenUnderestimationGlobalVulnerability2015, Brienen et al. 2015].

@@

The structure and microclimates of global forests are changing due to unprecedented disturbance rates.  Loss of canopy cover due to forest disturbances (e.g., deforestation, fire, and drought) is associated with reductions in canopy structural complexity, and altered microclimates and surface energy balance components (Zellweger et al. 2020; Stark et al. 2020 ). Reductions in canopy cover can lead to non-linear threshold responses, causing dramatic shifts from one forest state to another (e.g., transition from forest to savanna-like vegetation in tropical forest regions through "savannization"), with energy balance impacts and implications for forest-atmosphere interactions (Stark et al. 2020).
